Visualising metagenome differences

### 2023-11-22 analysis/GO_and_ortho.Rmd
make the setup blocks nicer
new Venn plots

```{r setup}
library(tidyverse)
library(patchwork)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

@@

```{r}
load('output/all_GO_plots.Rdata')
load('output/group_venn_image.Rdata')
load('output/seagrass_venn_image.Rdata')
```
```{r}
group_venn
```
```{r}
seagrass_venn
```

@@

>>>>

```{r setup, message=FALSE}
library(tidyverse)
library(patchwork)
library(wesanderson)
library(RColorBrewer)
library(UpSetR)
library(kableExtra)
library(ggVennDiagram)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

@@

```{r}
load('output/all_GO_plots.Rdata')
load('output/group_venn_image.Rdata')
load('output/seagrass_venn_image.Rdata')
```
```{r}
group_venn
```
```{r}
```
```{r}
seagrass_venn
```

@@

```{r}
a <- group_venn
b <- plot_list$GO_results_genes_in_Terrestrials_NOT_Aquatics_NOT_Seagrasses.csv
c <- seagrass_venn
d <- plot_list$GO_results_genes_in_all_seagrasses_vs_seagrassesBackground_intersect.csv
sum3 <- (a + c) / (b + d) +
   plot_annotation(tag_levels = 'A') +
  plot_layout(heights=c(1, 1.5))
sum3
cowplot::save_plot(sum3, filename = 'output/patchwork_terrestrials_seagrasses_gene_loss.png', base_height = 10)
```

### 2023-11-22 analysis/GOenrichment.Rmd
new Venn plots

```{r setup, message=FALSE}
library(topGO)
library(tidyverse)
library(rvest)
library(wesanderson)
library(httr)
library(stringi)
library(ggrepel)
library(eulerr)
library(UpSetR)
library(rrvgo)
library(kableExtra)
library(patchwork)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

@@

@@

for (f in list.files('./output/', pattern='*.csv')){
  filename <- paste('./output/', f, sep='')
  go_and_pvalues <- readChar(filename, file.info(filename)$size)
  go_and_pvalues <- gsub(',', ' ', go_and_pvalues)

  httr::POST(
    url = "http://revigo.irb.hr/StartJob.aspx",
    body = list(
      cutoff = "0.7",
      valueType = "pvalue",
     # speciesTaxon = "4577", # zea mays
     #speciesTaxon = '39947', # japonica
     speciesTaxon = '3702', # arabidopsis
      measure = "SIMREL",
      goList = go_and_pvalues
    ),
    # application/x-www-form-urlencoded
    encode = "form"
  ) -> res

  dat <- httr::content(res, encoding = "UTF-8")

  jobid <- jsonlite::fromJSON(dat,bigint_as_char=TRUE)$jobid

  # Check job status
  running <- "1"
  while (running != "0" ) {
      httr::POST(
        url = "http://revigo.irb.hr/QueryJobStatus.aspx",
        query = list( jobid = jobid )
      ) -> res2
      dat2 <- httr::content(res2, encoding = "UTF-8")
      running <- jsonlite::fromJSON(dat2)$running
      Sys.sleep(1)
  }

  # Fetch results
  httr::POST(
    url = "http://revigo.irb.hr/ExportJob.aspx",
    query = list(
      jobid = jobid, 
      namespace = "1",
      type = "CSVTable"
    )
  ) -> res3

  dat3 <- httr::content(res3, encoding = "UTF-8")

  dat3 <- stri_replace_all_fixed(dat3, "\r", "")
  # Now we have a csv table in a string!

  # read_csv does not like the ', ', it wants ','
  dat <- read_csv(gsub(', ', ',', dat3), show_col_types = FALSE)

  # do we even have results?
  if(nrow(dat) == 0){next}
  results_list[[f]] <- dat

@@

```{r warning=FALSE, message=FALSE}
# lots of warnings and messages here so I'm hiding these
plot_list <- list()
for (f in names(results_list)) {
  dat <- results_list[[f]]
  if(nrow(dat) == 0) {next}

  names(dat) <- c("term_ID","description","frequency", 'plot_X', 'plot_Y', 'log_size', 'value', 'uniqueness', 'dispensability', 'eliminated', 'representative')
  one.data <- dat
  if (one.data[1,] == '<html>') {next} # some datasets result in error
  one.data <- one.data [(one.data$plot_X != "null" & one.data$plot_Y != "null") & (! is.na(one.data$frequency)) & (one.data$value != 'null'), ];
  one.data <- suppressMessages(type_convert(one.data)) # guess data types, but also: shush
  top_ten_values <- one.data %>% arrange(value) %>% head(8) %>% pull(term_ID)

  one.data <- one.data %>% mutate(description2 = case_when(
    term_ID %in% top_ten_values ~ str_wrap(description, width=10), 
    TRUE ~ ''))
  p1 <-
    ggplot(data = one.data, aes(plot_X, plot_Y, label = description2)) +
    geom_point(aes(colour = value, size = log_size), alpha = I(0.6)) +
    scale_size_area() +
    scale_colour_gradientn(colours = rev(wes_palette("Zissou1", 100, type = "continuous")), 
                           limits = c(min(one.data$value), 0)) +
    geom_point(
      aes(plot_X, plot_Y, size = log_size),
      shape = 21,
      fill = "transparent",
      colour = I (alpha ("black", 0.6))
    ) +
    scale_size_area() +
    geom_label_repel(max.overlaps=15,
                     point.padding = 0,
                     min.segment.length = 0, # draw all line segments
                     aes(point.size=log_size), alpha=0.9) + 
    theme_minimal() +
    labs(color = 'log(p-value)',
         size = 'Frequency') +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    ) +
    NULL
    #ggtitle(f)

  plot_list[[f]] <- p1

@@

```{r results='asis'}
for( i in names(results_list)) {
  print(i)
  print(results_list[[i]] %>% dplyr::filter(Eliminated == FALSE) %>% dplyr::select(TermID, Name, Value) %>% arrange(Value) %>% kbl() %>%  kable_styling())
  cat("\n")
}

@@

```{r}
a <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$Name,
          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$Name,
          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$Name,
          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$Name)

a_go_ids <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$TermID,
          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$TermID,
          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$TermID,
          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$TermID)
```

@@

```{r}
a_no_ara <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$Name,
          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name)
```

>>>>

```{r setup, message=FALSE}
library(topGO)
library(tidyverse)
library(rvest)
library(rrvgo)
library(wesanderson)
library(httr)
library(stringi)
library(ggrepel)
library(eulerr)
library(UpSetR)
library(rrvgo)
library(kableExtra)
library(patchwork)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

@@

```{r}

@@

for (f in list.files('./output/', pattern='*.csv')){
  if(str_detect(f, '.png')) {
    next
  }
  filename <- paste('./output/', f, sep='')
  go_and_pvalues <- readChar(filename, file.info(filename)$size)
  go_and_pvalues <- gsub(',', ' ', go_and_pvalues)
  temp <- read_delim(go_and_pvalues,show_col_types = FALSE)
  if(nrow(temp) < 2) {next}
  simMatrix <- calculateSimMatrix(temp$name,
                                  orgdb="org.At.tair.db",
                                  ont="BP",
                                  method="Rel")

  tmp <- temp$value
  names(tmp) <- temp$name
  reducedTerms <- reduceSimMatrix(simMatrix,
                                  tmp,
                                  threshold=0.7,
                                  orgdb="org.At.tair.db")
  results_list[[f]] <- list('simMatrix'=simMatrix, 'reducedTerms'=reducedTerms)

  #httr::POST(
  #  url = "http://revigo.irb.hr/StartJob",
  #  body = list(
  #    cutoff = "0.7",
  #    valueType = "pvalue",
  #   # speciesTaxon = "4577", # zea mays
  #   #speciesTaxon = '39947', # japonica
  #    speciesTaxon = '3702', # arabidopsis
  #    measure = "SIMREL",
  #    goList = go_and_pvalues
  #  ),
  #  # application/x-www-form-urlencoded
  #  encode = "form"
  #) -> res
# 
#   dat <- httr::content(res, encoding = "UTF-8")
#   
#   if(typeof(dat)!="list")
#   {
#     jobid <- jsonlite::fromJSON(dat,bigint_as_char=TRUE)$jobid
#   } else {
#     jobid <- dat$jobid
#   }  
#   # Check job status
#   running <- "1"
#   while (running != "0" ) {
#       httr::GET(
#         url = "http://revigo.irb.hr/QueryJob",
#         query = list( jobid = jobid, type="jstatus" )
#       ) -> res2
#       dat2 <- httr::content(res2, encoding = "UTF-8")
#       # fix when httr::content automatically converts json to list
#       if(typeof(dat2)!="list")
#       {
#         running <- jsonlite::fromJSON(dat2)$running
#       } else {
#         running <- dat2$running
#       }
#       Sys.sleep(1)
#   }
#   
#   # Fetch results
#   httr::GET(
#     url = "http://revigo.irb.hr/QueryJob",
#     query = list(
#       jobid = jobid, 
#       namespace = "1",
#       type = "table"
#     )
#   ) -> res3
#   dat3 <- httr::content(res3, encoding = "UTF-8")
#   
#   dat3 <- stri_replace_all_fixed(dat3, "\r", "")
#   # Now we have a csv table in a string!
# 
#   # read_csv does not like the ', ', it wants ','
#   dat <- read_tsv(gsub(', ', ',', dat3), show_col_types = FALSE) 
#   
#   # do we even have results?
#   if(nrow(dat) == 0){next}
#   results_list[[f]] <- dat
#   
}
```

@@

```{r warning=FALSE, message=FALSE}
# lots of warnings and messages here so I'm hiding these
plot_list <- list()
for (f in names(results_list)) {

  tmp <- results_list[[f]]
  simMatrix <- tmp[['simMatrix']]
  reducedTerms <- tmp[['reducedTerms']]
  if(nrow(reducedTerms) <= 2) {next}
  plot_list[[f]] <- scatterPlot(simMatrix, reducedTerms, labelSize=3)
  # old REVIGO code
  # dat <- results_list[[f]]
  # if(nrow(dat) == 0) {next}
  # 
  # names(dat) <- c("term_ID","description","frequency", 'plot_X', 'plot_Y', 'log_size', 'value', 'uniqueness')#, 'dispensability', 'eliminated', 'representative')
  # one.data <- dat
  # #if (one.data[1,] == '<html>') {next} # some datasets result in error
  # #one.data <- one.data [(one.data$plot_X != "null" & one.data$plot_Y != "null") & (! is.na(one.data$frequency)) & (one.data$value != 'null'), ];
  # #one.data <- suppressMessages(type_convert(one.data)) # guess data types, but also: shush
  # top_ten_values <- one.data %>% arrange(value) %>% head(8) %>% pull(term_ID) 
  # 
  # one.data <- one.data %>% mutate(description2 = case_when(
  #   term_ID %in% top_ten_values ~ str_wrap(description, width=10), 
  #   TRUE ~ ''))
  # p1 <-
  #   ggplot(data = one.data, aes(plot_X, plot_Y, label = description2)) +
  #   geom_point(aes(colour = value, size = log_size), alpha = I(0.6)) +
  #   scale_size_area() +
  #   scale_colour_gradientn(colours = rev(wes_palette("Zissou1", 100, type = "continuous")))+ #, 
  #                          #limits = c(min(one.data$value), 0)) +
  #   geom_point(
  #     aes(plot_X, plot_Y, size = log_size),
  #     shape = 21,
  #     fill = "transparent",
  #     colour = I (alpha ("black", 0.6))
  #   ) +
  #   scale_size_area() +
  #   geom_label_repel(max.overlaps=15,
  #                    point.padding = 0,
  #                    min.segment.length = 0, # draw all line segments
  #                    aes(point.size=log_size), alpha=0.9) + 
  #   theme_minimal() +
  #   labs(color = 'log(p-value)',
  #        size = 'Frequency') +
  #   theme(
  #     axis.text.x = element_blank(),
  #     axis.ticks.x = element_blank(),
  #     axis.text.y = element_blank(),
  #     axis.ticks.y = element_blank(),
  #     axis.title.x = element_blank(),
  #     axis.title.y = element_blank()
  #   ) +
  #   NULL
  #   #ggtitle(f)
  # 
  # plot_list[[f]] <- p1
  # 
}
```

@@

```{r results='asis'}
for( i in names(results_list)) {
  print(i)
  #print(results_list[[i]] %>% dplyr::select(TermID, Name, Value) %>% arrange(Value) %>% kbl() %>%  kable_styling())
  print(results_list[[i]]$reducedTerms %>% dplyr::select(go, parent, score) %>% arrange(score) %>% kbl() %>%  kable_styling())
  cat("\n")
}

@@

```{r}
a <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$reducedTerms$term,
          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$reducedTerms$term,
          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$reducedTerms$term,
          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$reducedTerms$term)

a_go_ids <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$reducedTerms$go,
          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$reducedTerms$go,
          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$reducedTerms$go,
          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$reducedTerms$go)
```

@@

```{r}
a_no_ara <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$reducedTerms$term,
          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$reducedTerms$term,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$reducedTerms$term,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$reducedTerms$term)
```

### 2023-11-22 analysis/metagenome.Rmd
new Venn plots

```{r}
genessplit |> filter(str_detect(gene, 'nifH') | str_detect(gene, 'acdS') | str_detect(gene, 'sctC')) |> 
  mutate(gene_class = case_when(str_detect(gene, 'nifH') ~ '*nifH*',
                                str_detect(gene, 'acdS') ~ '*acdS*',
                                str_detect(gene, 'sctC') ~ '*sctC*')) |>
  #group_by(Sample_species, Tissue, Sample_replicate, MAG) |> 
  #summarise(gene_class = paste0(sort(unique(gene_class)), collapse = ' & ')) |>
  group_by(Sample_species, Tissue, Sample_replicate) |> 
  count(gene_class) |> 
  mutate(Sample_species = trunc_str(Sample_species)) |> 
  mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
  ggplot(aes(x=factor(Tissue, levels=rev(c('Shoot/Leaf', 'Shoot/Leaf Biofilm', 'Rhizome', 'Root'))), y=n, fill=Sample_species)) + 
  geom_boxplot(outlier.shape = NA) +
  facet_wrap(~factor(gene_class, levels=c('*acdS*', '*nifH*', '*sctC*')), scales = 'free_x') + 
  geom_point(position = position_jitterdodge(), aes(color=Sample_species)) +
  coord_flip() +
  theme_minimal() +
  labs(color='Species', fill='Species') +
  theme(axis.text.y =  element_markdown(),legend.text = element_markdown(),
        strip.text.x = element_markdown()) +
  ylab('Count of MAGs') + xlab('Species')
```
I like this one more. The differences are stronger.

>>>>

```{r}
genessplit |> filter(str_detect(gene, 'nifH') | str_detect(gene, 'acdS') | str_detect(gene, 'sctC')) |> 
  mutate(gene_class = case_when(str_detect(gene, 'nifH') ~ '*nifH*',
                                str_detect(gene, 'acdS') ~ '*acdS*',
                                str_detect(gene, 'sctC') ~ '*sctC*')) |>
  group_by(Sample_species, Tissue, Sample_replicate) |> 
  count(gene_class) |> 
  mutate(Sample_species = trunc_str(Sample_species)) |> 
  mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
  ggplot(aes(x=factor(Tissue, levels=rev(c('Shoot/Leaf', 'Shoot/Leaf Biofilm', 'Rhizome', 'Root'))), y=n, fill=Sample_species)) + 
  geom_boxplot(outlier.shape = NA) +
  facet_wrap(~factor(gene_class, levels=c('*acdS*', '*nifH*', '*sctC*')), scales = 'free_x') + 
  geom_point(position = position_jitterdodge(), aes(color=Sample_species)) +
  coord_flip() +
  theme_minimal() +
  labs(color='Host species', fill='Host species') +
  theme(axis.text.y = element_markdown(),
        legend.text = element_markdown(),
        strip.text.x = element_markdown()) +
  ylab('Count of MAGs') + 
  xlab('Tissue')
```
I like this one more. The differences are stronger.
